### 2023-02-03 products/manuscript/Manuscript.qmd
Part 1 of MADA Project

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

**PART 1**

**DATA SOURCE AND DATA DESCRIPTION**

The data for the project is extracted from the National Health Interview Survey (NHIS). The NHIS survey collects information related to health, health care access, health behaviors of the civilian and non-institutionalized U.S. population and makes digital data files available for years 1963 to present. Users can create their own costumed NHIS data extracts for analysis from the given website: <https://nhis.ipums.org/nhis-action/variables/group>.

List of variables extracted in the dataset to be used for the study:

1.  Demographics: age, sex, race/ethnicity, marital status, veteran status, family

2.  Socio-economic status: education, family income, occupation

3.  General health: Health status, BMI

4.  Women health: pregnancy status, hormone replacement therapy status, use of birth control pills/implants/shots

5.  Cancer testing: mammogram screening

6.  Cancer family history: family members with cancer

7.  Health conditions: cancer, diabetes, high cholesterol

8.  Health behaviors: Alcohol consumption, smoking and pregnancy, physical activity

9.  Health insurance

10. Medical care and access

The extracted data set contains the above listed variables information for years 2000, 2005, 2010, and 2015. As most data were missing for most of the other years, I decided to go with the selected years (2000, 2005, 2010, and 2015) as there were most data points available for those years and 5-years difference comparison might give great insights at the trends of mammogram screening over years. The extracted data set has 393,032 rows and 138 columns.

**QUESTIONS/HYPOTHESIS TO BE ADDRESSED**

The main research question of this study is: Is there any association between mammogram screening for breast cancer and socio-economic status of people?

**METHOD OF ANALYSIS**

Multiple regressions models with robust standard errors will be used for examining the association between mammogram screening and socio-economic status of people adjusting for age, sex, race/ethnicity, family history with cancer, and other control variables. Predictor variables will be various socio-economic determinant variables while the outcome variable will be mammogram screening for breast cancer.

**---------------------------------------------------------------------------------------------------------------------------------**

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

*Provide enough background on your topic that others can understand the why and how of your analysis*

## Description of data and data source

*Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section.*

## Questions/Hypotheses to be addressed

*State the research questions you plan to answer with this analysis.*

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.*

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References