### 2023-02-24 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
Part 2 of MADA project

---
title: "Codes_for_EDA"
author: "Shiwani Sapkota"
date: "2023-02-24"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(gmodels) #for ch-sq freuency table
library(vcd) # for three way frequency table
library(gtsummary)
```

```{r}
# Reading the cleaned csv file
data_clean <- readRDS("mammogram_ses_clean.rds")
head(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
gg_miss_var(data_clean_main)

# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

CREATING DEMOGRAPHIC CHARACTERISTICS TABLE

```{r}
# Lets look at the distribution of hipanic and non-hispanic women during the given years
CrossTable(data_clean_main$year, data_clean_main$hispanic_status)

# Lets look at the mammogram status among hipanic and non-hispanic women during the given years
CrossTable(data_clean_main$mammogram_status, data_clean_main$hispanic_status)

# Lets look at the number of mammogram status of women aged 40-74 during the years
CrossTable(data_clean_main$year, data_clean_main$mammogram_status)

# Lets look only at the socio-economic characteristics of women aged 40-64 by hispanic race status for different years
# Lets look at the age distribution among hispanic and non-hipanic women during the given years
year_age_hisp <- data_clean_main %>% select(year, age, hispanic_status)
mytable1 <- xtabs(~year+age+hispanic_status, data = year_age_hisp)
ftable(mytable1)
ftable(addmargins(prop.table(mytable1, c(1, 2)), 3))*100

# Lets look at the mammogram distribution among hispanic and non-hipanic women during the given years
year_mammogram_hisp <- data_clean_main %>% select(year, mammogram_status, hispanic_status)
mytable2 <- xtabs(~year+mammogram_status+hispanic_status, data = year_mammogram_hisp)
ftable(mytable2)
ftable(addmargins(prop.table(mytable2, c(1, 2)), 3))*100

# Lets look at the income distribution among hispanic and non-hipanic women during the given years
year_income_hisp <- data_clean_main %>% select(year, income, hispanic_status)
mytable3 <- xtabs(~year+income+hispanic_status, data = year_income_hisp)
ftable(mytable3)
ftable(addmargins(prop.table(mytable3, c(1, 2)), 3))*100

# Lets look at the education status among hispanic and non-hipanic women during the given years
year_education_hisp <- data_clean_main %>% select(year, education, hispanic_status)
mytable4 <- xtabs(~year+education+hispanic_status, data = year_education_hisp)
ftable(mytable4)
ftable(addmargins(prop.table(mytable4, c(1, 2)), 3))*100

# Lets look at the marital status among hispanic and non-hipanic women during the given years
year_marital_hisp <- data_clean_main %>% select(year, marital_status, hispanic_status)
mytable5 <- xtabs(~year+marital_status+hispanic_status, data = year_marital_hisp)
ftable(mytable5)
ftable(addmargins(prop.table(mytable5, c(1, 2)), 3))*100

# Lets look at the region of residence among hispanic and non-hipanic women during the given years
year_residence_hisp <- data_clean_main %>% select(year, region_residence, hispanic_status)
mytable6 <- xtabs(~year+region_residence+hispanic_status, data = year_residence_hisp)
ftable(mytable6)
ftable(addmargins(prop.table(mytable6, c(1, 2)), 3))*100

# Lets look at the insurance status among hispanic and non-hipanic women during the given years
year_insurance_hisp <- data_clean_main %>% select(year, insurance_status, hispanic_status)
mytable7 <- xtabs(~year+insurance_status+hispanic_status, data = year_insurance_hisp)
ftable(mytable7)
ftable(addmargins(prop.table(mytable7, c(1, 2)), 3))*100

# Lets look at the usual medical care status among hispanic and non-hipanic women during the given years
year_usualmedicalcare_hisp <- data_clean_main %>% select(year, usual_medicalcare_status, hispanic_status)
mytable8 <- xtabs(~year+usual_medicalcare_status+hispanic_status, data = year_usualmedicalcare_hisp)
ftable(mytable8)
ftable(addmargins(prop.table(mytable8, c(1, 2)), 3))*100

# Lets look at the smoking status among hispanic and non-hipanic women during the given years
year_smoking_hisp <- data_clean_main %>% select(year, smoking_status, hispanic_status)
mytable9 <- xtabs(~year+smoking_status+hispanic_status, data = year_smoking_hisp)
ftable(mytable9)
ftable(addmargins(prop.table(mytable9, c(1, 2)), 3))*100

# Lets look at the alcohol status among hispanic and non-hipanic women during the given years
year_alcohol_hisp <- data_clean_main %>% select(year, alcohol_status, hispanic_status)
mytable10 <- xtabs(~year+alcohol_status+hispanic_status, data = year_alcohol_hisp)
ftable(mytable10)
ftable(addmargins(prop.table(mytable10, c(1, 2)), 3))*100

# Lets look at the diabetes status among hispanic and non-hipanic women during the given years
year_diabetes_hisp <- data_clean_main %>% select(year, diabetes_status, hispanic_status)
mytable11 <- xtabs(~year+diabetes_status+hispanic_status, data = year_diabetes_hisp)
ftable(mytable11)
ftable(addmargins(prop.table(mytable11, c(1, 2)), 3))*100
```

CREATING SOME FIGURES

```{r}

```