### 2023-02-24 code/processing_cleaning_code/Codes_for_cleaning.Rmd

---
title: "Codes_for_cleaning"
author: "Shiwani Sapkota"
date: "2023-02-24"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r}
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
```

LOADING DATA FILE

```{r}
# I will be working on rds file so loading and looking at the raw rds data file
setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
data_main <- readRDS("mammogram_ses1.rds")
```

LOOKING AT THE DATA

```{r}
# Looking at the data
glimpse(data_main)
summary(data_main)
skim(data_main)
```

DATA CLEANING

```{r}
# Converting the data file to data frame and converting all columns to lowercase
data_main_df <- data.frame(data_main)
names(data_main_df) <- tolower(names(data_main_df))

# Lets look at the columns to check if our above code worked
colnames(data_main_df)

# Restricting our study to only females i.e. 2
data_main_df <- data_main_df[(which(data_main_df$sex == 2)), ]

# Checking if our above code worked and if there are any missing values in `sex` variable
unique(data_main_df$sex)
dim(data_main_df)
sum(is.na(data_main_df$sex))

# Restricting our study to only females who are 40-74 years old
unique(data_main_df$age)
data_main_df <- subset(data_main_df, data_main_df$age >= 40 & data_main_df$age <= 74)

# Checking if our above code worked and if there are any missing values in `age` variable
unique(data_main_df$age)
dim(data_main_df)
sum(is.na(data_main_df$age))

# Categorizing the `age` variable into 7 categories each with 5 years gap
head(data_main_df$age, n = 5)
data_main_df$age[data_main_df$age >= 40 & data_main_df$age <= 44] <- 1
data_main_df$age[data_main_df$age >= 45 & data_main_df$age <= 49] <- 2
data_main_df$age[data_main_df$age >= 50 & data_main_df$age <= 54] <- 3
data_main_df$age[data_main_df$age >= 55 & data_main_df$age <= 59] <- 4
data_main_df$age[data_main_df$age >= 60 & data_main_df$age <= 64] <- 5
data_main_df$age[data_main_df$age >= 65 & data_main_df$age <= 69] <- 6
data_main_df$age[data_main_df$age >= 70 & data_main_df$age <= 74] <- 7

# Checking to see if our `age` variable was categorized correctly as we wanted
head(data_main_df$age, n = 5)
unique(data_main_df$age)

# Looking at `hispyn` variable (hispanic ethnicity, dichotomous: 1 as non-hispanic and 2 as hipanic ethnicity)
unique(data_main_df$hispyn)
data_main_df <- data_main_df %>% 
  rename("hispanic_status" = "hispyn")
sum(is.na(data_main_df$hispanic_status))
dim(data_main_df)

# Looking at the `pooryn` above or below poverty threshold variable
unique(data_main_df$pooryn)

# Renaming `pooryn` variable to `income`
data_main_df <- data_main_df %>% 
  rename("income" = "pooryn")

# Filtering only at or above poverty threshold: 1 and below poverty threshold:2
data_main_df <- filter(data_main_df, income %in% c(1, 2))
unique(data_main_df$income)
dim(data_main_df)
sum(is.na(data_main_df$income))

# Looking at the `educrec2` education attainment variable
unique(data_main_df$educrec2)

# Renaming `educrec2` variable to `education`
data_main_df <- data_main_df %>% 
  rename("education" = "educrec2")

# Filtering out the `education` variable contents
data_main_df <- filter(data_main_df, education %in% c(10, 20, 31, 32, 41, 42, 51, 54, 60))

# Recoding `education` variable
# education: 1:	Never attended/kindergarten only
#            2:	Less than high school
#            3:	High school
#            4:	Some college or Associate's degree
#            5:	Bachelor's degree and higher
data_main_df$education[data_main_df$education == 10] <- 1
data_main_df$education[data_main_df$education == 20 | data_main_df$education == 31 | data_main_df$education == 32 | data_main_df$education == 41] <- 2
data_main_df$education[data_main_df$education == 42] <- 3
data_main_df$education[data_main_df$education == 51] <- 4
data_main_df$education[data_main_df$education == 54 | data_main_df$education == 60] <- 5
sum(is.na(data_main_df$education))
dim(data_main_df)
unique(data_main_df$education)

# Looking at the `marstcohab` education attainment variable
unique(data_main_df$marstcohab)

# Renaming `marstcohab` variable to `marital_status`
data_main_df <- data_main_df %>% 
  rename("marital_status" = "marstcohab")

# Removing the unknown marital status
data_main_df <- filter(data_main_df, marital_status %in% c(1, 2, 4, 5, 6, 7, 8))
unique(data_main_df$marital_status)

# Recoding `marital_status` variable
# marital_status: 0:	Never married
#                 1:	Married
#                 2:	Single (separated/widowed/divorced)
#                 3:	Living with partner
data_main_df$marital_status[data_main_df$marital_status == 8] <- 0
data_main_df$marital_status[data_main_df$marital_status == 1 | data_main_df$marital_status == 2] <- 1
data_main_df$marital_status[data_main_df$marital_status == 4 | data_main_df$marital_status == 5 | data_main_df$marital_status == 6] <- 2
data_main_df$marital_status[data_main_df$marital_status == 7] <- 3
unique(data_main_df$marital_status)
dim(data_main_df)

# Looking at the `hinotcove` health insurance coverage variable
unique(data_main_df$hinotcove)

# Renaming `hinotcove` variable to `insurance_status`
data_main_df <- data_main_df %>% 
  rename("insurance_status" = "hinotcove")

# Removing the unknown insurance status
data_main_df <- filter(data_main_df, insurance_status %in% c(1, 2))
unique(data_main_df$insurance_status)

# Recoding `insurance_status` variable
# insurance_status: 0:	No coverage
#                   1:	Coverage
data_main_df$insurance_status[data_main_df$insurance_status == 1] <- 0
data_main_df$insurance_status[data_main_df$insurance_status == 2] <- 1
sum(is.na(data_main_df$insurance_status))
unique(data_main_df$insurance_status)
dim(data_main_df)

# Looking at the `hiprivatee` private insurance variable
unique(data_main_df$hiprivatee)

# Renaming `hiprivatee` variable to `privateinsurance_status`
data_main_df <- data_main_df %>% 
  rename("privateinsurance_status" = "hiprivatee")

# Removing the unknown and no information private insurance status
# privateinsurance_status: 1: No
#                          2: Yes
data_main_df <- filter(data_main_df, privateinsurance_status %in% c(1, 2))
sum(is.na(data_main_df$privateinsurance_status))
unique(data_main_df$privateinsurance_status)

# Looking at the `himcaide` medicaid status variable
unique(data_main_df$himcaide)

# Renaming `himcaide` variable to `medicaid_status`
data_main_df <- data_main_df %>% 
  rename("medicaid_status" = "himcaide")

# Removing the unknown and no information medicaid status
# medicaid_status: 1: No
#                  2: Yes
data_main_df <- filter(data_main_df, medicaid_status %in% c(1, 2))
unique(data_main_df$medicaid_status)
sum(is.na(data_main_df$medicaid_status))

# Looking at the `himcaree` medicare status variable
unique(data_main_df$himcaree)

# Renaming `himcaree` variable to `medicare_status`
data_main_df <- data_main_df %>% 
  rename("medicare_status" = "himcaree")

# Removing no information medicare status
# medicare_status: 1: No
#                  2: Yes
data_main_df <- filter(data_main_df, medicare_status %in% c(1, 2))
unique(data_main_df$medicare_status)
sum(is.na(data_main_df$medicare_status))

# Looking at the `health` health status variable
unique(data_main_df$health)

# Renaming `health` variable to `health_status`
data_main_df <- data_main_df %>% 
  rename("health_status" = "health")

# Removing unknown health status and recoding the variable
# health_status: 1: Excellent/very good/good
#                2: Fair or poor
data_main_df <- filter(data_main_df, health_status %in% c(1, 2, 3, 4, 5))
data_main_df$health_status[data_main_df$health_status == 1 | data_main_df$health_status == 2 | data_main_df$health_status == 3] <- 1
data_main_df$health_status[data_main_df$health_status == 4 | data_main_df$marital_status == 5] <- 2
sum(is.na(data_main_df$health_status))
dim(data_main_df)

# Looking at the `mamev` ever had a mammogram variable
unique(data_main_df$mamev)

# Renaming `mamev` variable to `mammogram_status`
data_main_df <- data_main_df %>% 
  rename("mammogram_status" = "mamev")

# Removing unknown mammogram status and recoding the variable
# mammogram_status: 0: No
#                   1: Yes
data_main_df <- filter(data_main_df, mammogram_status %in% c(1, 2))
data_main_df$mammogram_status[data_main_df$mammogram_status == 1] <- 0
data_main_df$mammogram_status[data_main_df$mammogram_status == 2] <- 1
sum(is.na(data_main_df$mammogram_status))
dim(data_main_df)
unique(data_main_df$mammogram_status)

# Looking at the `cancerev` ever told had cancer variable
unique(data_main_df$cancerev)

# Renaming `cancerev` variable to `cancer_status`
data_main_df <- data_main_df %>% 
  rename("cancer_status" = "cancerev")

# Removing unknown cancer status and recoding the variable
# cancer_status: 0: No
#                1: Yes
data_main_df <- filter(data_main_df, cancer_status %in% c(1, 2))
unique(data_main_df$cancer_status)
data_main_df$cancer_status[data_main_df$cancer_status == 1] <- 0
data_main_df$cancer_status[data_main_df$cancer_status == 2] <- 1
sum(is.na(data_main_df$cancer_status))
dim(data_main_df)

# Limiting our study to only non-cancer people
data_main_df <- data_main_df[(which(data_main_df$cancer_status == 0)), ]
unique(data_main_df$cancer_status)
dim(data_main_df)

# Looking at the `smokestatus2` Cigarette smoking recode 2: Current detailed/former/never variable
unique(data_main_df$smokestatus2)

# Renaming `smokestatus2r` variable to `smoking_status`
data_main_df <- data_main_df %>% 
  rename("smoking_status" = "smokestatus2")

# Removing unknown smoking status and recoding the variable
# smoking_status: 0: Never
#                 1: Current
#                 2: Former
data_main_df <- filter(data_main_df, smoking_status %in% c(11, 12, 20, 30))
unique(data_main_df$smoking_status)
data_main_df$smoking_status[data_main_df$smoking_status == 30] <- 0
data_main_df$smoking_status[data_main_df$smoking_status == 11 | data_main_df$smoking_status == 12] <- 1
data_main_df$smoking_status[data_main_df$smoking_status == 20] <- 2
sum(is.na(data_main_df$smoking_status))
dim(data_main_df)

# Looking at the `alcstat1` Alcohol drinking status variable
unique(data_main_df$alcstat1)

# Renaming `alcstat1` variable to `alcohol_status`
data_main_df <- data_main_df %>% 
  rename("alcohol_status" = "alcstat1")

# Removing unknown alcohol status and recoding the variable
# smoking_status: 0: Never
#                 1: Current
#                 2: Former
data_main_df <- filter(data_main_df, alcohol_status %in% c(1:3))
unique(data_main_df$alcohol_status)
data_main_df$alcohol_status[data_main_df$alcohol_status == 1] <- 0
data_main_df$alcohol_status[data_main_df$alcohol_status == 3] <- 1
data_main_df$alcohol_status[data_main_df$alcohol_status == 2] <- 2
sum(is.na(data_main_df$alcohol_status))
dim(data_main_df)

# Looking at the `usualpl` Has usual place for medical care variable
unique(data_main_df$usualpl)

# Renaming `usualpl` variable to `usual_medicalcare_status`
data_main_df <- data_main_df %>% 
  rename("usual_medicalcare_status" = "usualpl")

# Removing unknown usual medicalcare status and recoding the variable
# usual_medicalcare_status: 0: No place
#                           1: Usual place
data_main_df <- filter(data_main_df, usual_medicalcare_status %in% c(1:3))
unique(data_main_df$usual_medicalcare_status)
data_main_df$usual_medicalcare_status[data_main_df$usual_medicalcare_status == 1] <- 0
data_main_df$usual_medicalcare_status[data_main_df$usual_medicalcare_status == 2 | data_main_df$usual_medicalcare_status == 3] <- 1
sum(is.na(data_main_df$usual_medicalcare_status))
dim(data_main_df)

# Looking at the `region` Region of residence variable
unique(data_main_df$region)

# Renaming `region` variable to `region_residence`
# region_residence: 1: Northeast
#                   2: North Central/Midwest
#                   3: South
#                   4: West
data_main_df <- data_main_df %>% 
  rename("region_residence" = "region")
sum(is.na(data_main_df$region_residence))

# Looking at the `diabeticev` Ever told had diabetes variable
unique(data_main_df$diabeticev)

# Renaming `diabeticev` variable to `diabetes_status`
data_main_df <-  data_main_df %>% 
  rename("diabetes_status" = "diabeticev")

# Removing unknown diabetes status and recoding the variable
# diabetes_status: 0: No
#                  1: Yes
data_main_df <- filter(data_main_df, diabetes_status %in% c(1, 2, 3))
data_main_df$diabetes_status[data_main_df$diabetes_status == 1] <- 0
data_main_df$diabetes_status[data_main_df$diabetes_status == 2 | data_main_df$diabetes_status == 3] <- 1
sum(is.na(data_main_df$diabetes_status))
dim(data_main_df)
```

SAVING THE CLEANED DATA FILE AS A CSV and RDS FILE

```{r}
# Commenting this out so that it does not get created many times
# data_main_clean <- subset(data_main_df, select = c("year", "strata", "psu", "perweight", "sampweight", "sex", "age", "hispanic_status", "income", "education", "marital_status", "region_residence", "insurance_status", "privateinsurance_status", "medicaid_status", "medicare_status", "health_status", "mammogram_status", "cancer_status", "usual_medicalcare_status", "smoking_status", "alcohol_status", "diabetes_status"))
# write_csv(data_main_clean, "mammogram_ses_clean.csv")
# saveRDS(data_main_clean, "mammogram_ses_clean.rds")
```