### 2023-02-25 code/analysis_code/exploratory_analysis/Codes_for_exploratory_analysis.Rmd
Project Part 2 Updates

---
title: "Codes_for_EDA"
author: "Shiwani Sapkota"
date: "2023-02-24"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for ch-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned csv file
setwd('/Users/ShiwaniSapkota/Desktop/UGA Semester 2 (Spring 2023) courses/EPID 8060E/Project/Raw Data from NHIS')
data_clean <- readRDS("mammogram_ses_clean.rds")
head(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
gg_miss_var(data_clean_main)

# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

CREATING DEMOGRAPHIC CHARACTERISTICS TABLE

```{r}
# Recoding variables to work in making table
data_clean_maint <- data_clean_main
data_clean_maint$hispanic_status <- recode_factor(data_clean_maint$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')
data_clean_maint$mammogram_status <- recode_factor(data_clean_maint$mammogram_status, '0' = 'No', '1' = 'Yes')
data_clean_maint$age <- recode_factor(data_clean_maint$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
data_clean_maint$income <- recode_factor(data_clean_maint$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
data_clean_maint$education <- recode_factor(data_clean_maint$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
data_clean_maint$marital_status <- recode_factor(data_clean_maint$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
data_clean_maint$region_residence <- recode_factor(data_clean_maint$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
data_clean_maint$insurance_status <- recode_factor(data_clean_maint$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
data_clean_maint$health_status <- recode_factor(data_clean_maint$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
data_clean_maint$usual_medicalcare_status <- recode_factor(data_clean_maint$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
data_clean_maint$smoking_status <- recode_factor(data_clean_maint$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$alcohol_status <- recode_factor(data_clean_maint$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_maint$diabetes_status <- recode_factor(data_clean_maint$diabetes_status, '0' = 'No', '1' = 'Yes')

# Lets look at the distribution of hipanic and non-hispanic women during the given years
CrossTable(data_clean_maint$year, data_clean_maint$hispanic_status)

# Lets look at the mammogram status among hipanic and non-hispanic women during the given years
CrossTable(data_clean_maint$mammogram_status, data_clean_maint$hispanic_status)

# Lets look at the number of mammogram status of women aged 40-74 during the years
CrossTable(data_clean_maint$year, data_clean_maint$mammogram_status)

# Lets look only at the socio-economic characteristics of women aged 40-64 by hispanic race status for different years
# Lets look at the age distribution among hispanic and non-hipanic women during the given years
year_age_hisp <- data_clean_maint %>% select(year, age, hispanic_status)
mytable1 <- xtabs(~year+age+hispanic_status, data = year_age_hisp)
ftable(mytable1)
ftable(addmargins(prop.table(mytable1, c(1, 2)), 3))*100

# Lets look at the mammogram distribution among hispanic and non-hipanic women during the given years
year_mammogram_hisp <- data_clean_maint %>% select(year, mammogram_status, hispanic_status)
mytable2 <- xtabs(~year+mammogram_status+hispanic_status, data = year_mammogram_hisp)
ftable(mytable2)
ftable(addmargins(prop.table(mytable2, c(1, 2)), 3))*100

# Lets look at the income distribution among hispanic and non-hipanic women during the given years
year_income_hisp <- data_clean_maint %>% select(year, income, hispanic_status)
mytable3 <- xtabs(~year+income+hispanic_status, data = year_income_hisp)
ftable(mytable3)
ftable(addmargins(prop.table(mytable3, c(1, 2)), 3))*100

# Lets look at the education status among hispanic and non-hipanic women during the given years
year_education_hisp <- data_clean_maint %>% select(year, education, hispanic_status)
mytable4 <- xtabs(~year+education+hispanic_status, data = year_education_hisp)
ftable(mytable4)
ftable(addmargins(prop.table(mytable4, c(1, 2)), 3))*100

# Lets look at the marital status among hispanic and non-hipanic women during the given years
year_marital_hisp <- data_clean_maint %>% select(year, marital_status, hispanic_status)
mytable5 <- xtabs(~year+marital_status+hispanic_status, data = year_marital_hisp)
ftable(mytable5)
ftable(addmargins(prop.table(mytable5, c(1, 2)), 3))*100

# Lets look at the region of residence among hispanic and non-hipanic women during the given years
year_residence_hisp <- data_clean_maint %>% select(year, region_residence, hispanic_status)
mytable6 <- xtabs(~year+region_residence+hispanic_status, data = year_residence_hisp)
ftable(mytable6)
ftable(addmargins(prop.table(mytable6, c(1, 2)), 3))*100

# Lets look at the insurance status among hispanic and non-hipanic women during the given years
year_insurance_hisp <- data_clean_maint %>% select(year, insurance_status, hispanic_status)
mytable7 <- xtabs(~year+insurance_status+hispanic_status, data = year_insurance_hisp)
ftable(mytable7)
ftable(addmargins(prop.table(mytable7, c(1, 2)), 3))*100

# Lets look at the usual medical care status among hispanic and non-hipanic women during the given years
year_usualmedicalcare_hisp <- data_clean_maint %>% select(year, usual_medicalcare_status, hispanic_status)
mytable8 <- xtabs(~year+usual_medicalcare_status+hispanic_status, data = year_usualmedicalcare_hisp)
ftable(mytable8)
ftable(addmargins(prop.table(mytable8, c(1, 2)), 3))*100

# Lets look at the smoking status among hispanic and non-hipanic women during the given years
year_smoking_hisp <- data_clean_maint %>% select(year, smoking_status, hispanic_status)
mytable9 <- xtabs(~year+smoking_status+hispanic_status, data = year_smoking_hisp)
ftable(mytable9)
ftable(addmargins(prop.table(mytable9, c(1, 2)), 3))*100

# Lets look at the alcohol status among hispanic and non-hipanic women during the given years
year_alcohol_hisp <- data_clean_maint %>% select(year, alcohol_status, hispanic_status)
mytable10 <- xtabs(~year+alcohol_status+hispanic_status, data = year_alcohol_hisp)
ftable(mytable10)
ftable(addmargins(prop.table(mytable10, c(1, 2)), 3))*100

# Lets look at the diabetes status among hispanic and non-hipanic women during the given years
year_diabetes_hisp <- data_clean_maint %>% select(year, diabetes_status, hispanic_status)
mytable11 <- xtabs(~year+diabetes_status+hispanic_status, data = year_diabetes_hisp)
ftable(mytable11)
ftable(addmargins(prop.table(mytable11, c(1, 2)), 3))*100

# Lets create a single complete demographic characteristics table by year and hispanic_status of women
table1 <- data_clean_maint %>% 
    select(mammogram_status, age, income, education, marital_status, region_residence, insurance_status, health_status, usual_medicalcare_status, smoking_status, alcohol_status, diabetes_status, year, hispanic_status) %>%
    tbl_strata(
      strata = year,
      .tbl_fun = 
        ~ .x %>%
        tbl_summary(
        missing = "no",
        by = c("hispanic_status"),
        statistic = all_categorical() ~ "{n} ({p}%)"))
table1
```

CREATING FIGURE

```{r}
# Summarizing mammogram_status and hispanic_status in the given years (Type I)
data_clean_main1 <- data_clean_main
data_clean_main1$hispanic_status <- recode_factor(data_clean_main1$hispanic_status, '1' = 'Non-hispanic Female', '2' = 'Hispanic Female')

# Lets save this figure
# png("figure1_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
ggplot(data = data_clean_main1, aes(x = mammogram_status)) +
  geom_histogram(fill = "white", colour = "red", stat = "count", position = position_dodge()) +
  facet_grid(hispanic_status ~ year) +
  scale_x_discrete(labels = c("No", "Yes")) +
  theme_bw()
# dev.off()

# Summarizing mammogram_status and hispanic_status in the given years (Type II)
# png("figure1optional_mammogram_ses.png", width = 10, height = 10, units = "in", res = 300)
ggplot(data_clean_main1, aes(x = mammogram_status))+
  geom_bar(aes(fill = hispanic_status), stat = "count", color = "white",
    position = position_dodge(0.9)) +
  facet_wrap(~year) +
  scale_x_discrete(labels = c("No", "Yes"))
# dev.off()
```