### 2023-03-20 code/analysis_code/statistical_analysis/Codes_for_statistical_analysis.Rmd
Updates addition

---
title: "Codes_for_statistical_analysis"
author: "Shiwani Sapkota"
date: "2023-03-16"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
```

FITTING LOGISTIC REGRESSION

Since our outcome variable `mammogram_status` is of binary type, we will be using bivariate and multivariate logistic regression methods for model fittings.

```{r}
# Setting up the logistic regression
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

BIVARIATE LOGISTIC REGRESSION

```{r}
# Using `mammogram_status` as the outcome and `year` as the predictor
log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
tidy(log_fit_year)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `age` as the predictor
log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
tidy(log_fit_age)
```

```{r}
# Using `mammogram_status` as the outcome and `income` as the predictor
log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
tidy(log_fit_income)
```

```{r}
# Using `mammogram_status` as the outcome and `education` as the predictor
log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)
tidy(log_fit_education)
```

```{r}
# Using `mammogram_status` as the outcome and `marital_status` as the predictor
log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)
tidy(log_fit_maritalstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `region_residence` as the predictor
log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)
tidy(log_fit_regionresidence)
```

```{r}
# Using `mammogram_status` as the outcome and `insurance_status` as the predictor
log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)
tidy(log_fit_insurancestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `health_status` as the predictor
log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)
tidy(log_fit_healthstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `usual_medicalcare_status` as the predictor
log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)
tidy(log_fit_usualmedicalcarestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `smoking_status` as the predictor
log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)
tidy(log_fit_smokingstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `alcohol_status` as the predictor
log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)
tidy(log_fit_alcoholstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `diabetes_status` as the predictor
log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)
tidy(log_fit_diabetesstatus)
```

MULTIVARIATE LOGISTIC REGRESSION

Fitting full model

```{r}
# Using `mammogram_status` as the outcome and the remaining variables as the predictors
log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
table2 <- tidy(log_fit_fullmodel)
table2
saveRDS(table2, file = (here("results", "statistical_analysis_result", "full_model.rds")))
```

Fitting model using some predictors

```{r}
# Using `mammogram_status` as the outcome and age & hispanic_status as predictors
log_fit_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
table3 <- tidy(log_fit_age_hispanic)
saveRDS(table3, file = (here("results", "statistical_analysis_result", "age_hispanic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and demographic characteristics variables as predictors
log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
table4 <- tidy(log_fit_demographic)
saveRDS(table4, file = (here("results", "statistical_analysis_result", "demographic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and the medical characteristics as the predictors
log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
table5 <- tidy(log_fit_medical)
saveRDS(table5, file = (here("results", "statistical_analysis_result", "medical_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome, and smoking & alcohol characteristics as the predictors
log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
table6 <- tidy(log_fit_smoking_alcohol)
saveRDS(table6, file = (here("results", "statistical_analysis_result", "smoking_alcohol_model.rds")))
```