### 2023-04-07 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd
Part 4 Addition

---
title: "Model Fitting"
author: "Shiwani Sapkota"
date: "2023-03-16"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))
```

FITTING LOGISTIC REGRESSION

Since our outcome variable `mammogram_status` is of binary type, we will be using bivariate and multivariate logistic regression methods for model fittings.

```{r}
# Setting up the logistic regression
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

BIVARIATE LOGISTIC REGRESSION

```{r}
# Using `mammogram_status` as the outcome and `year` as the predictor
log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
log_fit_year

# Looking at the summary of log_fit_year
tidy(log_fit_year)
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
log_fit_hispanicstatus

# Looking at the summary of log_fit_hispanicstatus
tidy(log_fit_hispanicstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `age` as the predictor
log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
log_fit_age

# Looking at the summary of log_fit_age
tidy(log_fit_age)
```

```{r}
# Using `mammogram_status` as the outcome and `income` as the predictor
log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
log_fit_income

# Looking at the summary of log_fit_income
tidy(log_fit_income)
```

```{r}
# Using `mammogram_status` as the outcome and `education` as the predictor
log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)
log_fit_education

# Looking at the summary of log_fit_education
tidy(log_fit_education)
```

```{r}
# Using `mammogram_status` as the outcome and `marital_status` as the predictor
log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)
log_fit_maritalstatus

# Looking at the summary of log_fit_maritalstatus
tidy(log_fit_maritalstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `region_residence` as the predictor
log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)
log_fit_regionresidence

# Looking at the summary of log_fit_regionresidence
tidy(log_fit_regionresidence)
```

```{r}
# Using `mammogram_status` as the outcome and `insurance_status` as the predictor
log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)
log_fit_insurancestatus

# Looking at the summary of log_fit_insurancestatus
tidy(log_fit_insurancestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `health_status` as the predictor
log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)
log_fit_healthstatus

# Looking at the summary of log_fit_healthstatus
tidy(log_fit_healthstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `usual_medicalcare_status` as the predictor
log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)
log_fit_usualmedicalcarestatus

# Looking at the summary of log_fit_usualmedicalcarestatus
tidy(log_fit_usualmedicalcarestatus)
```

```{r}
# Using `mammogram_status` as the outcome and `smoking_status` as the predictor
log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)
log_fit_smokingstatus

# Looking at the summary of log_fit_smokingstatus
tidy(log_fit_smokingstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `alcohol_status` as the predictor
log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)
log_fit_alcoholstatus

# Looking at the summary of log_fit_alcoholstatus
tidy(log_fit_alcoholstatus)
```

```{r}
# Using `mammogram_status` as the outcome and `diabetes_status` as the predictor
log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)
log_fit_diabetesstatus

# Looking at the summary of log_fit_diabetesstatus
tidy(log_fit_diabetesstatus)
```

MULTIVARIATE LOGISTIC REGRESSION

Fitting full model

```{r}
# Using `mammogram_status` as the outcome and the remaining variables as the predictors
log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
log_fit_fullmodel

# Looking at the summary of log_fit_fullmodel
table2 <- tidy(log_fit_fullmodel)
table2
saveRDS(table2, file = (here("results", "statistical_analysis_result", "model_fitting_result", "full_model.rds")))
```

Fitting model using some predictors

```{r}
# Using `mammogram_status` as the outcome and year, age, & hispanic_status as predictors
log_fit_year_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
log_fit_year_age_hispanic

# Looking at the summary of log_fit_year_age_hispanic
table3 <- tidy(log_fit_year_age_hispanic)
saveRDS(table3, file = (here("results", "statistical_analysis_result", "model_fitting_result", "year_age_hispanic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and demographic characteristics variables as predictors
log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
log_fit_demographic

# Looking at the summary of log_fit_demographic
table4 <- tidy(log_fit_demographic)
saveRDS(table4, file = (here("results", "statistical_analysis_result", "model_fitting_result", "demographic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and the medical characteristics as the predictors
log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
log_fit_medical

# Looking at the summary of log_fit_medical
table5 <- tidy(log_fit_medical)
saveRDS(table5, file = (here("results", "statistical_analysis_result", "model_fitting_result", "medical_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome, and smoking & alcohol characteristics as the predictors
log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
log_fit_smoking_alcohol

# Looking at the summary of log_fit_smoking_alcohol
table6 <- tidy(log_fit_smoking_alcohol)
saveRDS(table6, file = (here("results", "statistical_analysis_result", "model_fitting_result", "smoking_alcohol_model.rds")))
```

MODEL PERFORMANCE COMPARISON

```{r}
# Comparing all the above models on the basis of model performance
model_performance <- compare_performance(log_fit_year, log_fit_hispanicstatus, log_fit_age, log_fit_income, log_fit_education, log_fit_maritalstatus, log_fit_regionresidence, log_fit_insurancestatus, log_fit_healthstatus, log_fit_usualmedicalcarestatus, log_fit_smokingstatus, log_fit_alcoholstatus, log_fit_diabetesstatus, log_fit_fullmodel, log_fit_year_age_hispanic, log_fit_demographic, log_fit_medical, log_fit_smoking_alcohol)
table7 <- model_performance

# Saving model_performance result
saveRDS(table7, file = (here("results", "statistical_analysis_result", "model_fitting_result", "model_performance_comparison.rds")))
```