### 2023-04-07 code/analysis_code/statistical_analysis/2_model_evaluation/model_evaluation.Rmd

---
title: "Model Evaluation"
author: "Shiwani Sapkota"
date: "2023-04-07"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
library(broom)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))

# Changing class of mammogram_status variable from integer to factor as we can't use mammogram_status (main outcome) as an integer while preparing recipe and training the model
class(data_clean_main$mammogram_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$year <- as.factor(data_clean_main$year)
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

Let's start working on improving our model fits.

MODEL EVALUATION

DATA SPLITTING

```{r}
# Setting seed
set.seed(123)

# Putting 3/4 of the data into the training set
data_split <- initial_split(data_clean_main, prop = 3/4)

# Creating data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)
```

**FULL MODEL: CREATING RECIPE THAT FITS A LOGISTIC MODEL USING  AS OUTCOME OF INTEREST AND ALL OTHER VARIABLES AS PREDICTORS**

```{r}
# Using mammogram status as a categorical outcome of interest and all other variables of interest as predictors
data_clean_main_rec <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data)

# Fitting the logistic model
data_clean_main_mod <- logistic_reg() %>% 
                     set_engine("glm")

# Modelling workflow for pairing model and recipe 
data_clean_main_wflow <- workflow() %>% 
  add_model(data_clean_main_mod) %>% 
  add_recipe(data_clean_main_rec)
data_clean_main_wflow

# Using the resulting predictors for preparing recipe and training the model
data_clean_main_fit <- 
data_clean_main_wflow %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
data_clean_main_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**FULL MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE}
# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
predict(data_clean_main_fit, test_data)

# Using argument() with the model plus test data for saving them together
data_clean_main_aug <- 
  augment(data_clean_main_fit, test_data)

data_clean_main_aug %>%
  select(mammogram_status, .pred_0, .pred_1)

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug %>% 
  roc_auc(truth = mammogram_status, .pred_0)
```

**ALTERNATIVE MODEL: USING MAMMOGRAM STATUS AS OUTCOME OF INTEREST AND HISPANIC STATUS AS MAIN PREDICTOR**

```{r}
# Using mammogram status as a categorical outcome of interest and hispanic status as main predictor
data_clean_main_rec2 <- recipe(mammogram_status ~ hispanic_status, data = train_data)

# Fitting the logistic model
data_clean_main_mod2 <- logistic_reg() %>% 
                    set_engine("glm")

# Modelling workflow for pairing model and recipe 
data_clean_main_wflow2 <- workflow() %>% 
  add_model(data_clean_main_mod2) %>% 
  add_recipe(data_clean_main_rec2)
data_clean_main_wflow2

# Using the resulting predictors for preparing recipe and training the model
data_clean_main_fit2 <- 
 data_clean_main_wflow2 %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
data_clean_main_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE}
# Using the trained workflow (data_clean_main_fit) to predict with the unseen test data
predict(data_clean_main_fit2, test_data)

# Using argument() with the model plus test data for saving them together
data_clean_main_aug2 <- 
  augment(data_clean_main_fit2, test_data)

data_clean_main_aug2 %>%
  select(mammogram_status, .pred_0, .pred_1)

# Creating ROC curve and piping to the autoplot() method
data_clean_main_aug2 %>% 
  roc_curve(truth = mammogram_status, .pred_0) %>% 
  autoplot()

# Using roc_auc() for estimating the area under the curve 
data_clean_main_aug2 %>% 
  roc_auc(truth = mammogram_status, .pred_0)
```

From the above result, we see that the full model (using mammogram status as the main outcome of interest and other variables as predictors) has ROC-AUC as 0.76 while the alternative model (using mammogram status as the main outcome of interest and Hispanic status as the main predictor) has ROC-AUC as 0.54. So, we can say that for mammogram status, the full model with all the predictors of interest is better than the alternative model having only Hispanic status as the main predictor.