### 2023-04-08 code/analysis_code/statistical_analysis/3_machine_learning/machine_learning.Rmd
Part 4 Additional Updates

---
title: "Model Selection"
author: "Shiwani Sapkota"
date: "2023-04-07"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE}
library(here)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(ranger)
library(rpart)
library(glmnet)
library(vip)
library(rpart.plot)
library(performance)
library(yardstick)
library(recipes)
```

LOADING DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))

# Changing variables as factors
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$year <- as.factor(data_clean_main$year)
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)
```

**DATA SETUP**

```{r}
# Setting the random seed to 123
set.seed(123)

# Splitting dataset into 70% training and 30% testing and using mammogram status outcome as stratification
data_split <- initial_split(data_clean_main, prop = 7/10, strata = mammogram_status)

# Creating data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)

# Creating 5-fold cross-validation, 5 times repeated
CV_fold_data <- vfold_cv(train_data, v = 5, repeats = 5, strata = mammogram_status)

# Creating a recipe
recipe_data <- recipe(mammogram_status ~ year + age + hispanic_status + income + education + marital_status + region_residence + insurance_status + health_status + usual_medicalcare_status + smoking_status + alcohol_status + diabetes_status, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())
```

**NULL MODEL PERFORMANCE**

```{r}
# Using train data
# Creating a recipe for null model
recipe_null_train <- recipe(mammogram_status ~ 1, data = train_data) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Creating a logistic model recipe
recipe_null_logistic <- logistic_reg() %>% 
  set_engine("glm") %>% 
  set_mode("classification")

# Creating workflow pairing model and recipe 
workflow_null_train <- workflow() %>% 
  add_model(recipe_null_logistic) %>% 
  add_recipe(recipe_null_train)

# Fitting null model with the folds created from train data 
train_null <- fit_resamples(workflow_null_train, resamples = CV_fold_data)

# Computing ROC AUC for train data
metrics_null_train <- collect_metrics(train_null)
metrics_null_train
```

From the null model, we got ROC AUC as 0.5.

DECISION TREE MODEL: MODEL TUNING AND FITTING

```{r}
# Tuning hyperparameters by creating model specification that identifies which hyperparameters we are planning to tune
tune_tree_model <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()) %>%
  set_engine("rpart") %>% 
  set_mode("classification")
tune_tree_model

# Creating a regular grid of values for using some convenience functions for each hyperparameter
tree_grid <- grid_regular(cost_complexity(), tree_depth(), levels = 5)
tree_grid %>% count(tree_depth)

# Tuning the workflow using model specification and recipe and model
tree_wf <- workflow() %>%
  add_model(tune_tree_model) %>%
  add_recipe(recipe_data)

# Tuning using cross-validation and the tune_grid() function
tree_res <- tree_wf %>% 
  tune_grid(resamples = CV_fold_data,
    grid = tree_grid)
```

DECISION TREE MODEL: MODEL EVALUATION

```{r}
# Plotting and saving the above results
decision_tree <- tree_res %>% autoplot()
decision_tree
ggsave(filename = here("results","statistical_analysis_result", "3_machine_learning_result", "figure2_decision_tree_model.png"), plot = decision_tree)

# Getting the best-fit model
tree_res %>%
  show_best()
best_tree <- tree_res %>%
  select_best(metric = "roc_auc")

# Getting the final workflow
final_wf <- tree_wf %>% 
  finalize_workflow(best_tree)

# Fitting to the training data with the final workflow
final_fit <- final_wf %>%
  fit(train_data)

# Plotting the final fit
rpart.plot(extract_fit_parsnip(final_fit)$fit, roundint = FALSE)

# Predicted outcomes
predicted_fit <- predict(final_fit, train_data)
```

From the decision-tree model, we got ROC AUC (Area under the receiver operating characteristic curve) as 0.7.

LASSO MODEL: MODEL TUNING AND FITTING

```{r}
# Building the model
model_lasso <- logistic_reg(penalty = tune(), mixture = 1) %>% set_engine("glmnet")

# We will be using the recipe (recipe_data) that we created above 
# Creating the workflow
wf_lasso <- workflow() %>% 
  add_model(model_lasso) %>% 
  add_recipe(recipe_data)

# Model tuning using grid
lasso_grid <- tibble(penalty = 10^seq(-4, -1, length.out = 30))
lasso_grid %>% top_n(-5)
lasso_grid %>% top_n(5)

# Tuning the model using tune_grid() function
res_lasso <- wf_lasso %>% tune_grid(resamples = CV_fold_data, grid = lasso_grid,
    control = control_grid(verbose = FALSE, save_pred = TRUE), metrics = NULL)
res_lasso %>% collect_metrics()
```

LASSO MODEL: MODEL EVALUATION

```{r}
# Plotting the above results
res_lasso %>% autoplot()

# Getting the best-fit model
res_lasso %>%
  show_best()
best_lasso <- res_lasso %>%
  select_best(metric = "roc_auc")

# Getting the final workflow
final_lasso_wf <- wf_lasso %>% 
  finalize_workflow(best_lasso)

# Fitting to the training data with the final workflow
final_lasso_fit <- final_lasso_wf %>% fit(train_data)

# Plotting and saving the final fit
plot_lasso <- extract_fit_engine(final_lasso_fit)
lasso_model_figure <- plot(plot_lasso, "lambda")
lasso_model_figure
ggsave(filename = here("results","statistical_analysis_result", "3_machine_learning_result", "figure3_lasso_model.png"), plot = lasso_model_figure)

# Predicted outcomes
predicted_lasso_fit <- predict(final_lasso_fit, train_data)

# Looking at the coefficients and penalty for each variable in Lasso model
summary(best_lasso)
coefficients <- final_lasso_fit %>% extract_fit_parsnip() %>% tidy()
final_lasso_fit %>% extract_fit_parsnip() %>% tidy()

# Saving the above coefficients and penalty data in a table
table10 <- final_lasso_fit
saveRDS(table10, file = (here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds")))
```

From the Lasso model, we got ROC AUC as 0.8.

Based on the above results, lasso model performed the best as it had the highest ROC AUC compared to decision tree model. Hence, lasso model was selected.

FINAL EVALUATION

```{r}
# Fitting the final lasso model on split data
final_lasso_test <- final_lasso_wf %>% last_fit(data_split) 
final_lasso_test %>% collect_metrics()
```

From the above final model, we got ROC AUC as 0.8.