### 2023-04-08 products/supplement/Supplementary_Material.qmd

---
title: "Supplement to Manuscript for MADA Project"
subtitle: "Supplement Materials"
author: Shiwani Sapkota
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../dataanalysis_template_references.bib
csl: ../../vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(knitr)
library(ipumsr)
library(readr)
library(skimr)
library(tidyverse)
library(dplyr)
library(survey)
library(naniar)
library(gmodels) #for ch-sq frequency table
library(vcd) # for three way frequency table
library(gtsummary)
library(openxlsx)
library(ggplot2)
library(ggthemes)
```

# Overview

In this supplement file, you can find details about the data file extraction and codes involved in the process of data extraction and data analysis.

# Code and file information

```{r}
# Skip this code chunk if you already have `mammogram_ses1.rds` raw data file
# NHIS provides .xml and .dat file while extracting the data from website
# Commenting these codes as they were used for data extraction and data preparation process which are needed to be done only once before preparing the data file

# Preparing and loading the raw data file and saving as .csv and .rds files
# ddi <- read_ipums_ddi("nhis_00003.xml")
# data <- read_ipums_micro(ddi)

# Saving the raw data file as .csv file and commenting out this code so it does not get created everytime we run this code chunk
# readr::write_csv(data, "mammogram_ses.csv")

# It seems the created csv file is very large (153 MB) and GitKraken won't like it so lets create an rds file so that could be of small size

# Saving the raw data file as .rds file and commenting out this code so it does not get created everytime we run this code chunk
# saveRDS(data, "mammogram_ses1.rds")

# From here, you can follow general steps involved in data analysis and the
# respective codes are documented in the respected folders of this project:
# 1. Data cleaning
# 2. Exploratory Data analysis
# 3. Main statistical analysis
```

## Additional result

```{r}
#| label: Additional Table 1
#| tbl-cap: "Fit table using mammogram status as the outcome and age & hispanic_status as predictors."
#| echo: FALSE
additionaltable1 = readRDS(here("results", "statistical_analysis_result", "age_hispanic_model.rds"))
knitr::kable(additionaltable1)
```

```{r}
#| label: Additional Table 2
#| tbl-cap: "Fit table using mammogram status as the outcome and demographic characteristics variables as predictors."
#| echo: FALSE
additionaltable2 = readRDS(here("results", "statistical_analysis_result", "demographic_model.rds"))
knitr::kable(additionaltable2)
```

```{r}
#| label: Additional Table 3
#| tbl-cap: "Fit table using mammogram status as the outcome and the medical characteristics as the predictors."
#| echo: FALSE
additionaltable3 = readRDS(here("results", "statistical_analysis_result", "medical_model.rds"))
knitr::kable(additionaltable3)
```

```{r}
#| label: Additional Table 4
#| tbl-cap: "Fit table using mammogram status as the outcome and smoking & alcohol characteristics as the predictors."
#| echo: FALSE
additionaltable4 = readRDS(here("results", "statistical_analysis_result", "smoking_alcohol_model.rds"))
knitr::kable(additionaltable4)
```

```{r}
#| label: Additional Table 5
#| tbl-cap: "Model fitting table using mammogram status as the outcome and demographic characteristics as the predictors."
#| echo: FALSE
additionaltable5 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds"))
knitr::kable(additionaltable5)
```

```{r}
#| label: Additional Table 6
#| tbl-cap: "Model fitting table using mammogram status as the outcome and medical characteristics as the predictors."
#| echo: FALSE
additionaltable6 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds"))
knitr::kable(additionaltable6)
```

```{r}
#| label: Additional Table 7
#| tbl-cap: "Model fitting table using mammogram status as the outcome and smoking and alcohol characteristics as the predictors."
#| echo: FALSE
additionaltable7 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds"))
knitr::kable(additionaltable7)
```

```{r}
#| label: Additional Table 8
#| tbl-cap: "Model fitting table using mammogram status as the outcome and year, age, and hispanic characteristics as the predictors."
#| echo: FALSE
additionaltable8 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds"))
knitr::kable(additionaltable8)
```

```{r}
#| label: Additional Table 9
#| tbl-cap: "Model performance comparison."
#| echo: FALSE
additionaltable9 = readRDS(here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds"))
knitr::kable(additionaltable9)
```

```{r}
#| label: Additional Table 10
#| tbl-cap: "Model coefficients table using mammogram status as the outcome and other variables of interests as the predictors using model evaluation."
#| echo: FALSE
additionaltable10 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_full_model.rds"))
knitr::kable(additionaltable10)
```

```{r}
#| label: Additional Table 11
#| tbl-cap: "Model coefficients table using mammogram status as the outcome and hispanic status as the main predictor using model evaluation."
#| echo: FALSE
additionaltable11 = readRDS(here("results", "statistical_analysis_result", "3_machine_learning_result", "lasso_model_final_fit.rds"))
knitr::kable(additionaltable11)
```

```{r}
#| label: Additional Table 12
#| tbl-cap: "Lasso model final fit using machine learning approach."
#| echo: FALSE
additionaltable12 = readRDS(here("results", "statistical_analysis_result", "2_model_evaluation_result", "logistic_outcome_mainpredictor.rds"))
knitr::kable(additionaltable12)
```

```{r}
#| label: Additional Figure 1
#| fig-cap: "Decision tree model."
#| echo: FALSE
knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure2_decision_tree_model.png")
```

```{r}
#| label: Additional Figure 2
#| fig-cap: "Lasso model."
#| echo: FALSE
knitr::include_graphics("/results/statistical_analysis_result/3_machine_learning_result/figure3_lasso_model.png")
```

# Discussion

[**To be added]

# References

[**To be added]