### 2023-05-05 code/analysis_code/statistical_analysis/1_model_fitting/model_fitting.Rmd

---
title: "Model Fitting"
author: "Shiwani Sapkota"
date: "2023-03-16"
output: html_document
---

LOADING REQUIRED PACKAGES

```{r, warning = FALSE, message = FALSE}
# Loading required packages
library(readr)
library(tidyverse)
library(dplyr)
library(gtsummary)
library(ggplot2)
library(ggthemes)
library(here)
library(tidymodels)
library(performance)
library(vip)
library(gt)
```

LOADING AND LOOKING AT THE DATA

```{r}
# Reading the cleaned rds file
data_clean <- readRDS(here("data", "processed_cleaned_data", "mammogram_ses_clean.rds"))
glimpse(data_clean)

# Creating a new object to work with for the analysis
data_clean_main <- data_clean

# Making sure there are no missing values in any variables
any(is.na(data_clean_main))

# Lets convert the variables to factor variables just to make sure that they are not treated as continuous variables in our analysis
data_clean_main$age <- as.factor(data_clean_main$age)
data_clean_main$hispanic_status <- as.factor(data_clean_main$hispanic_status)
data_clean_main$income <- as.factor(data_clean_main$income)
data_clean_main$education <- as.factor(data_clean_main$education)
data_clean_main$marital_status <- as.factor(data_clean_main$marital_status)
data_clean_main$region_residence <- as.factor(data_clean_main$region_residence)
data_clean_main$insurance_status <- as.factor(data_clean_main$insurance_status)
data_clean_main$privateinsurance_status <- as.factor(data_clean_main$privateinsurance_status)
data_clean_main$medicaid_status <- as.factor(data_clean_main$medicaid_status)
data_clean_main$medicare_status <- as.factor(data_clean_main$medicare_status)
data_clean_main$health_status <- as.factor(data_clean_main$health_status)
data_clean_main$mammogram_status <- as.factor(data_clean_main$mammogram_status)
data_clean_main$usual_medicalcare_status <- as.factor(data_clean_main$usual_medicalcare_status)
data_clean_main$smoking_status <- as.factor(data_clean_main$smoking_status)
data_clean_main$alcohol_status <- as.factor(data_clean_main$alcohol_status)
data_clean_main$diabetes_status <- as.factor(data_clean_main$diabetes_status)

# Recoding the variables
data_clean_main$hispanic_status <- recode_factor(data_clean_main$hispanic_status, '1' = 'Non-Hispanic', '2' = 'Hispanic')
data_clean_main$mammogram_status <- recode_factor(data_clean_main$mammogram_status, '0' = 'No', '1' = 'Yes')
data_clean_main$age <- recode_factor(data_clean_main$age, '1' = '40-44 years', '2' = '45-49 years', '3' = '50-54 years', '4' = '55-59 years', '5' = '60-64 years', '6' = '65-69 years', '7' = '70-74 years')
data_clean_main$income <- recode_factor(data_clean_main$income, '1' = 'At or above poverty threshold', '2' = 'Below poverty threshold')
data_clean_main$education <- recode_factor(data_clean_main$education, '1' = 'Never attended/kindergarten only', '2' = 'Less than high school', '3' = 'High school', '4' = 'Some college or Associate degree', '5' = 'Bachelor degree and higher')
data_clean_main$marital_status <- recode_factor(data_clean_main$marital_status, '0' = 'Never married', '1' = 'Married', '2' = 'Single (separated/widowed/divorced)', '3' = 'Living with partner')
data_clean_main$region_residence <- recode_factor(data_clean_main$region_residence, '1' = 'Northeast', '2' = 'North Central/Midwest', '3' = 'South', '4' = 'West')
data_clean_main$insurance_status <- recode_factor(data_clean_main$insurance_status, '0' = 'No Coverage', '1' = 'Coverage')
data_clean_main$health_status <- recode_factor(data_clean_main$health_status, '1' = 'Excellent/very good/good', '2' = 'Fair/poor')
data_clean_main$usual_medicalcare_status <- recode_factor(data_clean_main$usual_medicalcare_status, '0' = 'No place', '1' = 'Usual place')
data_clean_main$smoking_status <- recode_factor(data_clean_main$smoking_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_main$alcohol_status <- recode_factor(data_clean_main$alcohol_status, '0' = 'Never', '1' = 'Current', '2' = 'Former')
data_clean_main$diabetes_status <- recode_factor(data_clean_main$diabetes_status, '0' = 'No', '1' = 'Yes')

# Limiting only 3 decimal places
options(digits = 3)
```

FITTING LOGISTIC REGRESSION

Since our outcome variable `mammogram_status` is of binary type, we will be using bivariate and multivariate logistic regression methods for model fittings.

```{r}
# Setting up the logistic regression
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

BIVARIATE LOGISTIC REGRESSION

```{r}
# Using `mammogram_status` as the outcome and `year` as the predictor
log_fit_year <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year), data = data_clean_main)
tidy(log_fit_year)

# Looking at the summary of log_fit_year
bivariate_year <- tidy(log_fit_year) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_year
saveRDS(bivariate_year, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_year.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `hispanic_status` as the predictor
log_fit_hispanicstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_hispanicstatus)

# Looking and saving the summary of log_fit_hispanicstatus
bivariate_hispanic <- tidy(log_fit_hispanicstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_hispanic
saveRDS(bivariate_hispanic, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_hispanic.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `age` as the predictor
log_fit_age <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age), data = data_clean_main)
tidy(log_fit_age)

# Looking at the summary of log_fit_age
tidy(log_fit_age)
bivariate_age <- tidy(log_fit_age) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_age
saveRDS(bivariate_age, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_age.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `income` as the predictor
log_fit_income <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(income), data = data_clean_main)
tidy(log_fit_income)

# Looking at the summary of log_fit_income
tidy(log_fit_income)
bivariate_income <- tidy(log_fit_income) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_income
saveRDS(bivariate_income, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_income.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `education` as the predictor
log_fit_education <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(education), data = data_clean_main)

# Looking at the summary of log_fit_education
tidy(log_fit_education)
bivariate_education <- tidy(log_fit_education) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_education
saveRDS(bivariate_education, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_education.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `marital_status` as the predictor
log_fit_maritalstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(marital_status), data = data_clean_main)

# Looking at the summary of log_fit_maritalstatus
tidy(log_fit_maritalstatus)
bivariate_maritalstatus <- tidy(log_fit_maritalstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_maritalstatus
saveRDS(bivariate_maritalstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_maritalstatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `region_residence` as the predictor
log_fit_regionresidence <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(region_residence), data = data_clean_main)

# Looking at the summary of log_fit_regionresidence
tidy(log_fit_regionresidence)
bivariate_regionresidence <- tidy(log_fit_regionresidence) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_regionresidence
saveRDS(bivariate_regionresidence, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_regionresidence.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `insurance_status` as the predictor
log_fit_insurancestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status), data = data_clean_main)

# Looking at the summary of log_fit_insurancestatus
tidy(log_fit_insurancestatus)
bivariate_insurancestatus <- tidy(log_fit_insurancestatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_insurancestatus
saveRDS(bivariate_insurancestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_insurancestatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `health_status` as the predictor
log_fit_healthstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(health_status), data = data_clean_main)

# Looking at the summary of log_fit_healthstatus
tidy(log_fit_healthstatus)
bivariate_healthstatus <- tidy(log_fit_healthstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_healthstatus
saveRDS(bivariate_healthstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_healthstatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `usual_medicalcare_status` as the predictor
log_fit_usualmedicalcarestatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(usual_medicalcare_status), data = data_clean_main)

# Looking at the summary of log_fit_usualmedicalcarestatus
tidy(log_fit_usualmedicalcarestatus)
bivariate_usualmedicalcarestatus <- tidy(log_fit_usualmedicalcarestatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_usualmedicalcarestatus
saveRDS(bivariate_usualmedicalcarestatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_usualmedicalcarestatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `smoking_status` as the predictor
log_fit_smokingstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status), data = data_clean_main)

# Looking at the summary of log_fit_smokingstatus
tidy(log_fit_smokingstatus)
bivariate_smokingstatus <- tidy(log_fit_smokingstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_smokingstatus
saveRDS(bivariate_smokingstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_smokingstatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `alcohol_status` as the predictor
log_fit_alcoholstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(alcohol_status), data = data_clean_main)

# Looking at the summary of log_fit_alcoholstatus
tidy(log_fit_alcoholstatus)
bivariate_alcoholstatus <- tidy(log_fit_alcoholstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_alcoholstatus
saveRDS(bivariate_alcoholstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_alcoholstatus.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and `diabetes_status` as the predictor
log_fit_diabetesstatus <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(diabetes_status), data = data_clean_main)

# Looking at the summary of log_fit_diabetesstatus
tidy(log_fit_diabetesstatus)
bivariate_diabetesstatus <- tidy(log_fit_diabetesstatus) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
bivariate_diabetesstatus
saveRDS(bivariate_diabetesstatus, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "bivariate_mammogram_diabetesstatus.rds")))
```

MULTIVARIATE LOGISTIC REGRESSION

Fitting full model

```{r}
# Using `mammogram_status` as the outcome and the remaining variables as the predictors
log_fit_fullmodel <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence) + as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(smoking_status) + as.factor(alcohol_status) + as.factor(diabetes_status), data = data_clean_main)
tidy(log_fit_fullmodel)

# Looking and saving the summary of log_fit_fullmodel
table2 <- tidy(log_fit_fullmodel) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
table2
saveRDS(table2, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "full_model.rds")))
```

Fitting model using some predictors

```{r}
# Using `mammogram_status` as the outcome and year, age, & hispanic_status as predictors
log_fit_year_age_hispanic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(year) + as.factor(age) + as.factor(hispanic_status), data = data_clean_main)
tidy(log_fit_year_age_hispanic)

# Looking at the summary of log_fit_year_age_hispanic
table3 <- tidy(log_fit_year_age_hispanic) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
table3
saveRDS(table3, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "year_age_hispanic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and demographic characteristics variables as predictors
log_fit_demographic <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(age) + as.factor(hispanic_status) + as.factor(income) + as.factor(education) + as.factor(marital_status) + as.factor(region_residence), data = data_clean_main)
tidy(log_fit_demographic)

# Looking at the summary of log_fit_demographic
table4 <- tidy(log_fit_demographic) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
table4
saveRDS(table4, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "demographic_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome and the medical characteristics as the predictors
log_fit_medical <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(insurance_status) + as.factor(health_status) + as.factor(usual_medicalcare_status) + as.factor(diabetes_status), data = data_clean_main)
tidy(log_fit_medical)

# Looking at the summary of log_fit_medical
table5 <- tidy(log_fit_medical) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
table5
saveRDS(table5, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "medical_model.rds")))
```

```{r}
# Using `mammogram_status` as the outcome, and smoking & alcohol characteristics as the predictors
log_fit_smoking_alcohol <- log_mod %>% fit(as.factor(mammogram_status) ~ as.factor(smoking_status) + as.factor(alcohol_status), data = data_clean_main)
tidy(log_fit_smoking_alcohol)

# Looking at the summary of log_fit_smoking_alcohol
table6 <- tidy(log_fit_smoking_alcohol) %>% 
  select(term, estimate, p.value) %>% 
  mutate('odds ratio' = exp(estimate)) %>% 
  select(term, 'odds ratio', p.value)
table6
saveRDS(table6, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "smoking_alcohol_model.rds")))
```

MODEL PERFORMANCE COMPARISON

```{r}
# Comparing all the above models on the basis of model performance
model_performance <- compare_performance(log_fit_year, log_fit_hispanicstatus, log_fit_age, log_fit_income, log_fit_education, log_fit_maritalstatus, log_fit_regionresidence, log_fit_insurancestatus, log_fit_healthstatus, log_fit_usualmedicalcarestatus, log_fit_smokingstatus, log_fit_alcoholstatus, log_fit_diabetesstatus, log_fit_fullmodel, log_fit_year_age_hispanic, log_fit_demographic, log_fit_medical, log_fit_smoking_alcohol)
table7 <- model_performance
table7

# Saving model_performance result
saveRDS(table7, file = (here("results", "statistical_analysis_result", "1_model_fitting_result", "model_performance_comparison.rds")))
```